---
title: "Deveopmental disorder phenotype"
output: html_document
---
```{r}
#install.packages(R.utils)

library(R.utils)

#install.packages(dplyr)
library(dplyr)

#install.packages(tidyr)
library(tidyr)

library(stringr)

#install.packages(tidyverse)
library(tidyverse)


```

#unzip file

```{r}
##gunzip("DDG2P_6_10_2023.csv.gz", remove=FALSE)

```

#read csv file

```{r}
DDP <- read.csv("DDG2P_6_10_2023.csv")
#DDP2 <- read.csv("DDG2P_6_10_2023.csv")
```

#label unknowns

```{r}
DDP$organ.specificity.list[DDP$organ.specificity.list == ''] <- 'Unknown'
#write.csv(DDP, "DDPwUnknown.csv")
```

#seperate out rows based on location/organs (will create duplicate rows of genes)(There are also duplicates before this anyway)

#creating counts table for numbers of genes in each location/organ

```{r}
DDPsep <- separate_longer_delim(DDP, organ.specificity.list,";")

##write.csv(DDPsep, "DDPsep.csv")

Counts <- as.data.frame(table(DDPsep$organ.specificity.list))
Counts <- Counts[-1,]

colnames(Counts)[1] <- "Organ/Location"

Counts <- Counts[order(Counts$Freq, decreasing = TRUE),]

```

```{r}
#write.csv(CountsV2, "CountsV2.csv")

```


```{r}
##x <- str_subset(DDPsep$organ.specificity.list, "Eye:")

##DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == x] <- "Eye"

```

##Collapsing suborgans into full organ category

```{r}
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Retina'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Lens'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Structure'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Physiology'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Cornea'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Connective tissue'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Optic nerve'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Vasculature'] <- 'Eye'
DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Eye: Vitreous'] <- 'Eye'

DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Endocrine'] <- 'Endocrine/Metabolic'

DDPsep$organ.specificity.list[DDPsep$organ.specificity.list == 'Peripheral nerves'] <- 'Spinal cord/Peripheral nerves'

#write.csv(DDPsep, "DDPsepV2.csv")
```

#creating counts table

```{r}
CountsV2 <- as.data.frame(table(DDPsep$organ.specificity.list))
CountsV2 <- CountsV2[-1,]

colnames(CountsV2)[1] <- "Organ/Location"

CountsV2 <- CountsV2[order(CountsV2$Freq, decreasing = TRUE),]
```

#creating list of just genes vs organs
```{r}
DDPsep <- read.csv("DDPsepV2.csv")

DDPLoc <- DDPsep[c('gene.symbol', 'organ.specificity.list')]

```
#Remove blank values

```{r}
#DDPLoc <- DDPLoc[!(is.na(DDPLoc$organ.specificity.list) | DDPLoc$organ.specificity.list==""), ]

```

###Making gene vs location table

###annotate blank spaces as unknown
```{r}
#DDP$organ.specificity.list[DDP$organ.specificity.list == ''] <- 'Unknown'

```

#merge duplicate listings


```{r}
#library(data.table)
DDPsep <- setDT(DDPsep)[,.(organ.specificity.list=paste(organ.specificity.list,collapse=",")),by="gene.symbol"]
#write.csv(DDP, "DDPnoDUP.csv")
```


#Get list of TRUE/FALSE values for each location; will be ordered by original order of genes in original csv

```{r}

Eye <- grepl("Eye", DDPsep$organ.specificity.list)
Brain <- grepl("Brain", DDPsep$organ.specificity.list)
Skeleton <- grepl("Skeleton", DDPsep$organ.specificity.list)
Endocrine <- grepl("Endocrine", DDPsep$organ.specificity.list)
Heart <- grepl("Heart", DDPsep$organ.specificity.list)
Face <- grepl("Face", DDPsep$organ.specificity.list)
Skin <- grepl("Skin", DDPsep$organ.specificity.list)
Kidney <- grepl("Kidney", DDPsep$organ.specificity.list)
GI <- grepl("GI", DDPsep$organ.specificity.list)
Ear <- grepl("Ear", DDPsep$organ.specificity.list)
Bone_Marrow <- grepl("Bone Marrow", DDPsep$organ.specificity.list)
Musculature <- grepl("Musculature", DDPsep$organ.specificity.list)
Genitalia <- grepl("Genitalia", DDPsep$organ.specificity.list)
Peripheral <- grepl("Peripheral", DDPsep$organ.specificity.list)
Teeth <- grepl("Teeth", DDPsep$organ.specificity.list)
Respiratory <- grepl("Respiratory", DDPsep$organ.specificity.list)
Hair <- grepl("Hair", DDPsep$organ.specificity.list)
Multisystem <- grepl("Multisystem", DDPsep$organ.specificity.list)
Cancer <- grepl("Cancer", DDPsep$organ.specificity.list)
Liver <- grepl("Liver", DDPsep$organ.specificity.list)
Lungs <- grepl("Lungs", DDPsep$organ.specificity.list)
Unknown <- grepl("Unknown", DDPsep$organ.specificity.list)



```


```{r}
sum(str_detect(Eye, "TRUE"))
sum(str_detect(Brain, "TRUE"))
sum(str_detect(Musculature, "TRUE"))

```

#Add TRUE/FALSE lists to create table

```{r}
LocGen <- subset(DDPsep, select = -c(2) )
LocGen$Eye <- Eye
LocGen$Brain_Cognition <- Brain
LocGen$Skeleton <- Skeleton
LocGen$Endocrine_Metabolic <- Endocrine
LocGen$Heart_Cardiovasculature_Lymphatic <- Heart
LocGen$Face <- Face
LocGen$Skin <- Skin
LocGen$Kidney_Renal_Tract <- Kidney
LocGen$GI_Tract <- GI
LocGen$Ear <- Ear
LocGen$Bone_Marrow_Immune <- Bone_Marrow
LocGen$Musculature <- Musculature
LocGen$Genitalia <- Genitalia
LocGen$Spinal_Cord_Peripheral_Nerves <- Peripheral
LocGen$Teeth_Dentation <- Teeth
LocGen$Respiratory_Tract <- Respiratory
LocGen$Hair_Nails <- Hair
LocGen$Multisystem <- Multisystem
LocGen$Cancer_Predisposition <- Cancer
LocGen$Liver <- Liver
LocGen$Lungs <- Lungs
LocGen$Unknown <- Unknown

```

#Find duplicates and remove


```{r}
#n_occur <- data.frame(table(LocGen$gene.symbol))

#LocGen = LocGen[!duplicated(LocGen$gene.symbol),]

LocGen <- LocGen %>% remove_rownames %>% column_to_rownames(var="gene.symbol") 


```

#number of duplicates
```{r}
n_occur[n_occur$Freq >1,]
```


```{r}
write.csv(LocGen, "gene_locationsV4.csv")
```

#Select rows(genes) based on location of body

```{r}
LocGen[LocGen$Ear == 'TRUE',]

```


```{r}
```


```{r}
```


```{r}
```




