---
title: "New Visualisation"
output: html_document
date: "2023-10-19"
---


```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(R.utils)
library(pheatmap)
library(UpSetR)
```



```{r}
DDP <- read.csv("gene_locationsV4.csv", row.names = 1)
# Load the PBMC dataset
pbmc.data <- Read10X(data.dir = "filtered_gene_bc_matrices/hg19")
# Initialize the Seurat object with the raw (non-normalized data).
pbmc <- CreateSeuratObject(counts = pbmc.data, project = "pbmc3k", min.cells = 3, min.features = 200)
pbmc[["percent.mt"]] <- PercentageFeatureSet(pbmc, pattern = "^MT-")
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
```


```{r}
DoHeatmap(pbmc)
```



```{r}
Difexp <- FetchData(pbmc, vars = all.genes)
Difexp
```

```{r}
DDPH <- sapply(DDP, as.integer)
DDPH <- as.data.frame(DDPH)
rownames(DDPH) <- rownames(DDP)
```

```{r}
upset(DDPH, nsets = 22,order.by = "freq")
```


```{r}
pbmc@assays$RNA@counts
head(pbmc[["RNA"]][[]])
```

#Finding similar genes between pbmc and ddd
```{r}

pbmcgen <- rownames(pbmc[["RNA"]][[]])
dddgen <- rownames(DDP)
length(intersect(pbmcgen,dddgen))

commongen <- intersect(pbmcgen,dddgen)
```

#Finding different genes

```{r}
diffgen <- setdiff(pbmcgen, dddgen)


```


#subsetting pbmc data based on these genes


```{r}

pbmc2 <- subset(pbmc,features=setdiff(rownames(pbmc),diffgen))

```
#Heatmap with reduced genes

```{r}
DoHeatmap(pbmc2)
```

```{r}
Skin <- DDP[DDP$Skin == "TRUE",]
Nervous <- DDP[DDP$Spinal_Cord_Peripheral_Nerves == "TRUE" | DDP$Brain_Cognition == "TRUE",]
Musco <- DDP[DDP$Skeleton == "TRUE" | DDP$Musculature,]
Circulatory <- DDP[DDP$Heart_Cardiovasculature_Lymphatic | DDP$Bone_Marrow_Immune,]

Skin <- rownames(Skin)
Nervous <- rownames(Nervous)
Musco <- rownames(Musco)
Circulatory <- rownames(Circulatory)
```

#Finding different genes between pbmc and ddd organ specific lists
```{r}
diffgenCirc <- setdiff(pbmcgen, Circulatory)
pbmcCirc <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenCirc))
DoHeatmap(pbmcCirc)
```

```{r}
diffgenSkin <- setdiff(pbmcgen, Skin)
pbmcSkin <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenSkin))
DoHeatmap(pbmcSkin)
```

```{r}
diffgenNervous <- setdiff(pbmcgen, Nervous)
pbmcNervous <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenNervous))
DoHeatmap(pbmcNervous)
```

```{r}
diffgenMusco <- setdiff(pbmcgen, Musco)
pbmcMusco <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenMusco))
DoHeatmap(pbmcMusco)
```

```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
pbmc <- RunUMAP(pbmc, dims = 1:10)
DimPlot(pbmc, reduction = "umap")
new.cluster.ids <- c("Lymphoid", "Myeloid", "Lymphoid", "Lymphoid", "Lymphoid", "Myeloid",
    "Lymphoid", "Myeloid", "Myeloid")
names(new.cluster.ids) <- levels(pbmc)
pbmc <- RenameIdents(pbmc, new.cluster.ids)
DimPlot(pbmc, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

-----------------------------------------------------------------------------------------------------------------------------------

#Re running heatmaps after clustering

#Finding similar genes between pbmc and ddd
```{r}

pbmcgen <- rownames(pbmc[["RNA"]][[]])
dddgen <- rownames(DDP)
length(intersect(pbmcgen,dddgen))

commongen <- intersect(pbmcgen,dddgen)
```

#Finding different genes

```{r}
diffgen <- setdiff(pbmcgen, dddgen)


```


#subsetting pbmc data based on these genes


```{r}

pbmc2 <- subset(pbmc,features=setdiff(rownames(pbmc),diffgen))

```
#Heatmap with reduced genes

```{r}
DoHeatmap(pbmc2)
```

```{r}
Skin <- DDP[DDP$Skin == "TRUE",]
Nervous <- DDP[DDP$Spinal_Cord_Peripheral_Nerves == "TRUE" | DDP$Brain_Cognition == "TRUE",]
Musco <- DDP[DDP$Skeleton == "TRUE" | DDP$Musculature,]
Circulatory <- DDP[DDP$Heart_Cardiovasculature_Lymphatic | DDP$Bone_Marrow_Immune,]

Skin <- rownames(Skin)
Nervous <- rownames(Nervous)
Musco <- rownames(Musco)
Circulatory <- rownames(Circulatory)
```

#Finding different genes between pbmc and ddd organ specific lists
```{r}
diffgenCirc <- setdiff(pbmcgen, Circulatory)
pbmcCirc <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenCirc))
DoHeatmap(pbmcCirc)
```

```{r}
diffgenSkin <- setdiff(pbmcgen, Skin)
pbmcSkin <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenSkin))
DoHeatmap(pbmcSkin)
```

```{r}
diffgenNervous <- setdiff(pbmcgen, Nervous)
pbmcNervous <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenNervous))
DoHeatmap(pbmcNervous)
```

```{r}
diffgenMusco <- setdiff(pbmcgen, Musco)
pbmcMusco <- subset(pbmc,features=setdiff(rownames(pbmc),diffgenMusco))
DoHeatmap(pbmcMusco)
```

